---
title: "p8105_hw3_ob2305"
author: "Olya Besedina"
data: "10/09/2019"
output: github_document
---
  
```{r setup, include=FALSE}
library(tidyverse)
library(viridis)

knitr::opts_chunk$set(
 	echo = TRUE,
 	warning = FALSE,
 	fig.width = 8, 
   fig.height = 6,
   out.width = "90%"
 )

options(
   ggplot2.continuous.colour = "viridis",
   ggplot2.continuous.fill = "viridis"
 )

 scale_colour_discrete = scale_colour_viridis_d
 scale_fill_discrete = scale_fill_viridis_d

 theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Problem 1

```{r include=FALSE}
library(p8105.datasets)
data("instacart")
```

* How many aisles are there, and which aisles are the most items ordered from?

There are 134 isles and most items are ordered from the fresh vegetables aisle. 

```{r}
instacart %>% 
  count(aisle)

instacart %>% 
  count(aisle, sort = TRUE) %>% 
  knitr::kable()
```

* Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

Fresh vegetables, fresh fruits, packaged vegetables fruits, yourt and packaged cheese are the top 5 ordered products. 

```{r}
instacart %>% 
  count(aisle, sort = TRUE) %>% 
  filter( n > 10000) %>% 
  ggplot(aes(x = n, y = aisle))+
  geom_point()
```

* Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

For “baking ingredients” 3 most popular itens are Organic Baby Spinach, Organic Raspberries and Organic Blueberries in descending order. For “packaged vegetables fruits” they are Light Brown Sugar, Pure Baking Soda and Cane Sugar. For “dog food care” the are  Snack Sticks Chicken & Rice Recipe Dog Treats, Organix Chicken & Brown Rice Recipe   and Small Dog Biscuits.

```{r}
bi = instacart %>%
  filter(aisle == "baking ingredients") %>%
  group_by(aisle, product_name) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n)) %>% 
  head(3)

dfc = instacart %>%
  filter(aisle == "dog food care") %>% 
  group_by(aisle, product_name) %>% 
  summarize(n = n()) %>%
  arrange(desc(n)) %>% 
  head(3)

pvf = instacart %>%
  filter(aisle == "packaged vegetables fruits") %>% 
  group_by(aisle, product_name) %>% 
  summarize(n = n()) %>%
  arrange(desc(n)) %>% 
  head(3) 

aisles_3 = 
  bind_rows(bi, dfc, pvf) %>%
  arrange(desc(n))
```

* Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

During the week coffee ice cream is ordered in the afternoon, while on saturday they are ordered at noon mostly. Apples are mostly ordered in the first half of the day and not later than 2 pm. 

```{r}
hour_of_day = instacart %>% 
  select(product_name, order_dow, order_hour_of_day) %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(
    mean_time_of_day = mean(order_hour_of_day))
  
hour_of_day %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_time_of_day)%>% 
  knitr::kable(digits = 1)

hour_of_day %>% 
  ggplot(aes(x = order_dow, y = mean_time_of_day, color = product_name))+
  geom_line()+
  labs(
    title = "Hour of order",
    x = "Day of order",
    y = "Mean time of day (Hour)")
```

mean of how many times some product was ordered 

# Problem 2

```{r}
library(p8105.datasets)
data("brfss_smart2010")

brfss = brfss_smart2010 %>%
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>%
  rename(
    state = locationabbr, 
    county = locationdesc)
```

* Check all possible responses

Responses include Excellent, Very good, Good, Fair anf Poor, so include all of them.

```{r}
brfss %>%
  count(response)
```


```{r}
brfss = brfss %>%
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>% 
  mutate(
    response = factor(response, levels = c("Excellent", "Very good", "Good", "Fair", "Poor"))) %>% 
  arrange(desc(response)) 

```

* In 2002, which states were observed at 7 or more locations? 

The "CT" "FL" "MA" "NC" "NJ" "PA" are the states which were observed at 7 or more locations in 2002. 

```{r}
first = brfss %>% 
  filter(year =="2002") %>% 
  group_by(state) %>%
  summarize(n_location = n_distinct(county)) %>% 
  filter(n_location >=7)

x = pull(first, state) 
```

* What about in 2010?

The "CA" "CO" "FL" "MA" "MD" "NC" "NE" "NJ" "NY" "OH" "PA" "SC" "TX" "WA" are the states which were observed at 7 or more locations in 2010.

```{r}
second = brfss %>% 
  filter(year =="2010") %>% 
  group_by(state) %>%
  summarize(n_location = n_distinct(county)) %>% 
  filter(n_location >=7)

y = pull(second, state)
```

* Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

Spaghetti plot depicts average data value for "Excellent" responses for time range of 2002 - 2010. Each state is depicted by different color. 

```{r}
brfss %>% 
  select(year, state, county, response, data_value) %>% 
  filter(response == "Excellent") %>% 
  group_by(response, year, state) %>% 
  summarize(
    ave_data_value = mean(data_value)
  ) %>% 
  ggplot(aes(x = year, y = ave_data_value, color = state))+
  geom_line()
```

* Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
brfss %>%
  select(state, year, data_value, response) %>% 
  filter(state =="NY", year %in% c(2006, 2010)) %>%
  ggplot(aes(x = data_value, fill = response))+
  geom_density(alpha = .5)+
  facet_grid(~year)
```

# Problem 3

```{r}
chf = read_csv(file = "./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(
    week_time = if_else(day == "Saturday" | day == "Sunday", "weekend", "weekday")) %>% 
  select(week, day_id, day,week_time, everything()) %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity_time",
    names_prefix = "activity_",
    values_to = "value") 
```

Tidied data set has `r ncol(chf)` variables with names:`r names(chf)` and `r nrow(chf)` data entries. 

```{r}
chf %>% head()
```

* Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

This spaghetti plot depicts total activity for the specific day over the course of 5 weeks. Monday of week 3 appears to be the day with the highest activity level, while saturday of week 5 has no activity at all. It appears that activity level overall started to go down after week 3.

```{r, eval=FALSE }
chf %>% 
  group_by(week, day) %>% 
  summarize(
    total_activity_per_day = sum(value)
  ) %>% 
  ggplot(aes(x = week, y = total_activity_per_day, color = day))+
  geom_line()+
  labs(
    title = "5 Weeks activity profile",
    x = "Week",
    y = "Activity value for the day")
```

* Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

This plot depicts activity value of the patient for every minute of the 5 weeks. It appears that Wednesdays have lower activity level compared to other days, while Fridays were the most active during the last week of trial.

```{r}
hour_data = 
  chf %>%
  group_by(week, day, day_id) %>%
  mutate(
   activity_time = as.numeric(activity_time),
    hour_activity = (activity_time)%/%60)

```

```{r}
hour_data %>% 
  filter(hour_activity != "24") %>% 
  ggplot(aes(x = activity_time, y = value, color = day, group = day_id))+
  geom_line(alpha = .5)+
  labs(
    title = "24 hour activity profile",
    x = "Minutes",
    y = "Activity Measure")
```


