---
title: "p8105_hw3_ob2305"
author: "Olya Besedina"
data: "10/09/2019"
output: github_document
---
  
```{r setup, include=FALSE}
library(tidyverse)
library(viridis)

knitr::opts_chunk$set(
 	echo = TRUE,
 	warning = FALSE,
 	fig.width = 8, 
   fig.height = 6,
   out.width = "90%"
 )

options(
   ggplot2.continuous.colour = "viridis",
   ggplot2.continuous.fill = "viridis"
 )

 scale_colour_discrete = scale_colour_viridis_d
 scale_fill_discrete = scale_fill_viridis_d

 theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Problem 1

```{r include=FALSE}
library(p8105.datasets)
data("instacart")
```

How many aisles are there, and which aisles are the most items ordered from?

There are 134 isles and most items are ordered from the fresh vegetables aisle. 

```{r}
instacart %>% 
  count(aisle)

instacart %>% 
  count(aisle, sort = TRUE) %>% 
  knitr::kable()

```

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}

instacart %>% 
  count(aisle, sort = TRUE) %>% 
  filter( n > 10000) %>% 
  ggplot(aes(x = n, y = aisle))+
  geom_point()
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
bi = instacart %>%
  filter(aisle == "baking ingredients") %>%
  group_by(aisle, product_name) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n)) %>% 
  head(3)

dfc = instacart %>%
  filter(aisle == "dog food care") %>% 
  group_by(aisle, product_name) %>% 
  summarize(n = n()) %>%
  arrange(desc(n)) %>% 
  head(3)

pvf = instacart %>%
  filter(aisle == "packaged vegetables fruits") %>% 
  group_by(aisle, product_name) %>% 
  summarize(n = n()) %>%
  arrange(desc(n)) %>% 
  head(3) 

aisles_3 = 
  bind_rows(bi, dfc, pvf) %>%
  arrange(desc(n))
```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).!!!!

```{r}
instacart %>% 
  select(product_name, order_dow, order_hour_of_day) %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(
    mean_time_of_day = mean(order_hour_of_day)
  ) %>% 
  view()
# produce a 2 x 7 table).!!!!
```

mean of how many times some product was ordered 

# Problem 2

```{r}
library(p8105.datasets)
data("brfss_smart2010")

brfss = brfss_smart2010 %>%
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>%
  rename(
    state = locationabbr, 
    county = locationdesc)
```

Check all possible responses

```{r}
brfss %>%
  count(response)
```

Responses include Excellent, Very good, Good, Fair anf Poor, so include all of them.

Clean and recode df

```{r}
brfss = brfss %>%
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>% 
  mutate(
    response = factor(response, levels = c("Excellent", "Very good", "Good", "Fair", "Poor"))) %>% 
  arrange(desc(response)) 

```

In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
first = brfss %>% 
  filter(year =="2002") %>% 
  group_by(state) %>%
  summarize(n_location = n_distinct(county)) %>% 
  filter(n_location >=7)
```

The `r pull(first, state)` are the states which were observed at 7 or more locations in 2002. 

What about in 2010?

```{r}
second = brfss %>% 
  filter(year =="2010") %>% 
  group_by(state) %>%
  summarize(n_location = n_distinct(county)) %>% 
  filter(n_location >=7)
```

Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
brfss %>% 
  select(year, state, county, response, data_value) %>% 
  filter(response == "Excellent") %>% 
  group_by(response, year, state) %>% 
  summarize(
    ave_data_value = mean(data_value)
  ) %>% 
  ggplot(aes(x = year, y = ave_data_value, color = state))+
  geom_line()
```

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
brfss %>%
  select(state, year, data_value, response) %>% 
  filter(state =="NY", year %in% c(2006, 2010)) %>%
  ggplot(aes(x = data_value, fill = response))+
  geom_density(alpha = .5)+
  facet_grid(~year)
```

# Problem 3

```{r}
chf = read_csv(file = "./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(
    week_time = if_else(day == "Saturday" | day == "Sunday", "weekend", "weekday")) %>% 
  select(week, day_id, day,week_time, everything()) %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity_time",
    names_prefix = "activity_",
    values_to = "value")

view(chf)

# describe data
```

Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r, eval=FALSE }
chf = chf %>% 
  group_by(week, day) %>% 
  summarize(
    total_activity = sum(value)
  ) %>% 
  ggplot(aes(x = week, y = total_activity, color = day))+
  geom_line()

# describe
```

Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
hour_data = 
  chf %>%
  group_by(week, day) %>%
  mutate(
   activity_time = as.numeric(activity_time)) %>% 
  mutate(
    hour_activity = (activity_time)%/%60) %>%
  group_by(hour_activity, day, week) %>%
  summarize(
    hour_value = (sum(value))%/%60)
```

```{r}
hour_data %>% 
  filter(hour_activity != "24") %>% 
  ggplot(aes(x = hour_value, y = hour_activity, color = day))+
  geom_line()
```






